Statically Typed String Sanitation Inside a Python

### 2014-08-06 sanitation-psp14/sanitation-psp14-short.tex
Updated PSP14 abstract

\documentclass{acm_proc_article-sp}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{afterpage}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\title{A Type System for String Sanitation\\Implemented Inside a Python}

%\authorinfo{~}{~}{~}
%dep_docclass\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
%dep_docclass           {School of Computer Science\\
%dep_docclass            Carnegie Mellon University}
%dep_docclass           {\{nathanfu, comar, aldrich\}@cs.cmu.edu}

\maketitle
\begin{abstract}
%Evidence suggests that programmers are reluctant to adopt new languages to
Security-oriented programming languages do not enjoy wide adoption. Conversely, library and framework-based approaches to security are pervasive. Unfortunately, libraries and frameworks are only effective solutions if they are implemented and used properly. In this paper, we propose that extensible programming languages provide a promising mechanism for increasing the amount of trust that can be place in libraries and frameworks. To demonstrate this approach, we present a conservative extension to the simply-typed lambda calculus for checking the correctness of input sanitation algorithms. We also demonstrate how this language's security guarantees are preserved under translation into an underlying language containing a regular expression library. We believe this approach -- complementing existing techniques with light-weight, composable type-based analyses -- constitutes a promising compromise between usability, potential for industry adoption, and theoretically grounded safety guarantees.
\end{abstract}

@@

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{r_1}{r_2}}$.
\end{lem}
\begin{proof}
The proof proceeds by structural induction on $r$.
The base case ($r=\alpha, \alpha \in \Sigma$) is trivial.
The choice and sequential cases require lemma 1, and prove a more general property that $\lsubst{r}{s_1}{s_2} \in \lang{r'}$ for some $r'$ equivalent to $\lreplace{r}{r_1}{r_2}$.
The closure case follows from the induction hypothesis applied to any $n$-unwinding.

>>>>

\documentclass[9pt]{sig-alternate}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{afterpage}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\title{Statically Typed String Sanitation Inside a Python}
\numberofauthors{3}
\author{
  \alignauthor
  Nathan Fulton
  \alignauthor
  Cyrus Omar
  \alignauthor
  Jonathan Aldrich
  \and
  \affaddr{Carnegie Mellon University}\\
  \affaddr{Pittsburgh, PA}\\
  \email{\{nathanfu, comar, aldrich\}@cs.cmu.edu}
}
%\authorinfo{~}{~}{~}
%\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
 %          {School of Computer Science\\
  %          Carnegie Mellon University
   %        \{nathanfu, comar, aldrich\}@cs.cmu.edu}

\maketitle
\begin{abstract}
%Evidence suggests that programmers are reluctant to adopt new languages to
Web applications must ultimately generate strings that contain commands to be consumed by systems like web browsers and database engines. If these strings are constructed from user input that has not been properly sanitized, this can expose costly injection vulnerabilities. %Structured frameworks for constructing commands centralizes control over string sanitization, but verifying that the framework itself sanitizes user input correctly can still be tricky.

In this paper, we introduce \emph{regular string types}, which classify strings known statically to be in a specified regular language. The language of a string is tracked by the type system through operations like concatenation, substitution and coercion, so regular string types can be used to implement, in essentially a conventional manner, the parts of a web application or application framework that constructs  commands. Simple type annotations at key points can be used to statically verify that sanitization has been performed correctly without introducing redundant run-time checks.

We take the position that to be practical, such a type system cannot require that programmers adopt a new programming language. Ideally, such a ``special-purpose'' type system would be distributed as a library that could safely be used together with other such type systems. We support this by specifying a sound translation to a simple language containing only strings and regular expressions, then discuss implementing the type system together with this translation as a library in \texttt{atlang}, an extensible static type system for Python (being developed by the authors).% represents the most sophisticated application of this library to date.
% Security-oriented programming languages do not enjoy wide adoption. Conversely, library and framework-based approaches to security are pervasive. Unfortunately, libraries and frameworks are only effective solutions if they are implemented and used properly. In this paper, we propose that extensible programming languages provide a promising mechanism for increasing the amount of trust that can be place in libraries and frameworks. To demonstrate this approach, we present a conservative extension to the simply-typed lambda calculus for checking the correctness of input sanitation algorithms. We also demonstrate how this language's security guarantees are preserved under translation into an underlying language containing a regular expression library. We believe this approach -- complementing existing techniques with light-weight, composable type-based analyses -- constitutes a promising compromise between usability, potential for industry adoption, and theoretically grounded safety guarantees.
\end{abstract}

@@

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{r_1}{r_2}}$.
\end{lem}
\begin{proof}
The proof proceeds by structural induction on $r$.
The base case ($r=\alpha, \alpha \in \Sigma$) is trivial.
The choice and sequential cases require lemma 1, and prove a more general property that $\lsubst{r}{s_1}{s_2} \in \lang{r'}$ for some $r'$ isomorphic to $\lreplace{r}{r_1}{r_2}$.
The closure case follows from the induction hypothesis applied to any $n$-unwinding.
